---
title: "Variant set (VS) analysis"
output: html_document
---


```{r, echo=F, warning=F, message=F}
rm(list=ls())
set.seed(123)
library(knitr)
library(RSQLite)
library(dplyr)
library(knitr)
library(kableExtra)
library(RColorBrewer)
library(gplots)
library(tidyverse)
library(gridExtra)
library(ggpubr)
#library("devtools")
#install.packages("Rtools")
#install_github('xinhe-lab/mirage')
library(mirage)
```



```{r burdern analysis-gene level, results='hide', echo=F, warning=F, message=F}
# read into gene set 
GeneDB=src_sqlite(path="C:\\Shengtong\\Research\\rare-var\\gene.list.db", create=F)
#GeneDB=src_sqlite(path="../../gene.list.db", create=F)
gene_cate1=data.frame(collect(tbl(GeneDB, "SFARI_HighConf")))
gene_cate2=data.frame(collect(tbl(GeneDB, "SFARI_StrongCand")))
gene_cate3=data.frame(collect(tbl(GeneDB, "SFARI_cate3_gene")))
gene_cate4=data.frame(collect(tbl(GeneDB, "SFARI_cate4_gene")))
gene_cate5=data.frame(collect(tbl(GeneDB, "SFARI_cate5_gene")))
gene_cate6=data.frame(collect(tbl(GeneDB, "SFARI_cate6_gene")))
gene_cateS=data.frame(collect(tbl(GeneDB, "SFARI_cateS_gene")))
TADAGene=data.frame(collect(tbl(GeneDB, "TADAGenelist")))
Qlessthan5percentgene=TADAGene$TadaName[TADAGene$qvalue.combined<0.05]
Qlessthan20percentgene=TADAGene$TadaName[TADAGene$qvalue.combined<0.2]
Qlessthan30percentgene=TADAGene$TadaName[TADAGene$qvalue.combined<0.3]
Qlessthan40percentgene=TADAGene$TadaName[TADAGene$qvalue.combined<0.4]
Qlessthan50percentgene=TADAGene$TadaName[TADAGene$qvalue.combined<0.5]
Qlargerthan90percentgene=TADAGene$TadaName[TADAGene$qvalue.combined>0.9]
###### ID gene ######
ID_Gene=as.character(data.frame(collect(tbl(GeneDB, "Pinto14AJHG_IDgene")))$GeneID)
#### high confidence gene 
High_Confidence_Gene=union(union(gene_cate1$GeneID, gene_cate2$GeneID),Qlessthan5percentgene)
#### moderate confidence gene 
Moderate_Confidence_Gene=setdiff(union(union(gene_cate3$GeneID, gene_cateS$GeneID), Qlessthan30percentgene),Qlessthan5percentgene)
#### psd gene 
purcell.genelist=data.frame(collect(tbl(GeneDB, "Purcell2014_genelist"))) ## PSD gene, SCZdenovo gene 
PSD_Gene=purcell.genelist$Gene_symbol[purcell.genelist$PSD=="Y"]
#### FMRP gene 
FMRP_Gene=purcell.genelist$Gene_symbol[purcell.genelist$FMRP.target=="Y"]
#### Autism KB gene 
ASD_KB_Gene=as.character(data.frame(collect(tbl(GeneDB, "AutismKB_gene")))$GeneID)
#### constraint gene 
Constraint_Gene=data.frame(collect(tbl(GeneDB, "Samocha_2014NG_constraintgene")))$gene
### RVIS gene 
RVIS.Allgene=data.frame(collect(tbl(GeneDB, "RVIS_gene")))
RVIS_Gene=RVIS.Allgene$GeneID[RVIS.Allgene$RVIS.percentile<5] # top 5% gene
#### Haplo insufficient gene 
Haploinsufficient_Gene=as.character(data.frame(collect(tbl(GeneDB, "Petrovski_plosgen_haploinsuff_gene")))$GeneID)
#### SCZ data 
SCZ_Gene=purcell.genelist$Gene_symbol
Gene_Set=c("ID gene","High", "Mod", "PSD", "FMRP", "AutismKB", "constraint gene", "RVIS", "Haploinsuff", "SCZ gene")
Gene_Set_List=list()
Gene_Set_List[[1]]=ID_Gene
Gene_Set_List[[2]]=High_Confidence_Gene
Gene_Set_List[[3]]=Moderate_Confidence_Gene
Gene_Set_List[[4]]=PSD_Gene
Gene_Set_List[[5]]=FMRP_Gene
Gene_Set_List[[6]]=ASD_KB_Gene
Gene_Set_List[[7]]=Constraint_Gene
Gene_Set_List[[8]]=RVIS_Gene
Gene_Set_List[[9]]=Haploinsufficient_Gene
Gene_Set_List[[10]]=SCZ_Gene
```


```{r, echo=F, cache=T, warning=F, message=F}
### old sample
########## read into data
#setwd("C:\\Shengtong\\Research\\rare-var\\rare-var-project\\analysis\\")
All.Anno.Data=as_tibble(read.table("../../AnnotatedTrans.txt", header=T))
N1=4315; N0=4315
All.Anno.Data[All.Anno.Data =="."] <- NA
All.Anno.Data$ExacAF[is.na(All.Anno.Data$ExacAF)]=0 # set AF of NA to zero 
Anno.Data=All.Anno.Data[which(All.Anno.Data$ExacAF<0.05 & All.Anno.Data$Annotation!="synonymous SNV"),] # use AF cutoff and exclude synonymous SNV
Anno.Data=All.Anno.Data[which(All.Anno.Data$Annotation!="synonymous SNV"),] # use AF cutoff and exclude synonymous SNV
```


```{r, echo=F, cache=T, warning=F, message=F}
### read into new sample 
ASC_new_sample1=as_tibble(read.table("C:\\Shengtong\\Research\\rare-var\\ASC\\ASC_B15_B16_raw_TDT_by_parent_2021-04-19.txt", header=T, fill=T))
ASC_new_sample2=as_tibble(read.table("C:\\Shengtong\\Research\\rare-var\\ASC\\ASC_v17_raw_TDT_by_parent_2021-05-10.txt", header=T, fill=T))
```


```{r, echo=F, eval=T}
## identify transmitted and non-transmitted variants 
#transmitted_variant=ASC_new_sample2%>%select(starts_with(c("t", "u")), "Gnomad_non_neuro_AF")
#kable(transmitted_variant[1:10,])%>%
#kable_styling() %>%
#scroll_box(height = "500px")
#write.csv(transmitted_variant, file="C:\\Shengtong\\Research\\rare-var\\ASC\\ASC_v17_raw_TDT_by_parent_2021-05-10_transmitted_variant.csv")

##### count transmitted variants and untransmitted in proband by summing over male, female, dad, mom and ind 
Transmitted_proband=ASC_new_sample2%>% select(starts_with("t_proband"))%>%  mutate(Transmitted_proband = select(., t_proband_male_dad:t_proband_female_ind) %>% rowSums())%>% select(Transmitted_proband) 
Untransmitted_proband=ASC_new_sample2%>% select(starts_with("u_proband"))%>%  mutate(Untransmitted_proband = select(., u_proband_male_dad:u_proband_female_ind) %>% rowSums())%>% select(Untransmitted_proband) 

##### count transmitted variants and untransmitted in proband from parents  by summing over male, female, dad, mom  
Transmitted_proband_parents=ASC_new_sample2%>% select(starts_with("t_proband"))%>%  mutate(Transmitted_proband_parents = select(., t_proband_male_dad, t_proband_male_mom, t_proband_female_dad, t_proband_female_mom) %>% rowSums())%>% select(Transmitted_proband_parents) 
Untransmitted_proband_parents=ASC_new_sample2%>% select(starts_with("u_proband"))%>%  mutate(Untransmitted_proband_parents= select(., u_proband_male_dad, u_proband_male_mom, u_proband_female_dad, u_proband_female_mom) %>% rowSums())%>% select(Untransmitted_proband_parents) 


##### add new columns of total transmitted and non-transmitted variants 
ASC_new_sample=ASC_new_sample2%>%mutate(Transmitted_proband=Transmitted_proband$Transmitted_proband, Untransmitted_proband=Untransmitted_proband$Untransmitted_proband, Transmitted_proband_parents=Transmitted_proband_parents$Transmitted_proband_parents, Untransmitted_proband_parents=Untransmitted_proband_parents$Untransmitted_proband_parents)

```


## Burden analysis for new sample 

### All rare variants 

* use filter `Gnomad_non_neuro_AF<0.05`

```{r, echo=F, message=F, warning=F}
num_proband=6026+1265 # male proband+female proband 
rare_variant_new_sample=ASC_new_sample %>% filter(Gnomad_non_neuro_AF<0.05) %>% select(Variant, Gene, Transmitted_proband, Untransmitted_proband)  # find syn variants   in new sample data with Gnomad AF threshold 

fisher_test=fisher.test(matrix(c(sum(rare_variant_new_sample$Transmitted_proband), num_proband, sum(rare_variant_new_sample$Untransmitted_proband), num_proband), nrow=2))
fisher_test_result_all=tibble(OR=fisher_test$estimate, lower=fisher_test$conf.int[1], upper=fisher_test$conf.int[2], var="all")
ggplot(fisher_test_result_all, aes(x=var, y=OR, fill=var)) + # draw the OR plot 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.2,
                 position=position_dodge(.9)) 
```


### Synonymous variants 


#### syn variants stratified by AF 

* `Gnomad_non_neuro_AF<0.05`

* Private variants are defined as variants with total allele  count equal to  1 in parents calculated as `Transmitted_proband_parents+Untransmitted_proband_parents=1`


```{r, echo=F}
rare_syn_variants_new_sample=ASC_new_sample %>% filter(Consequence=="synonymous_variant" & Gnomad_non_neuro_AF<0.05) %>% select(Variant, Gene, Gnomad_non_neuro_AF, Transmitted_proband, Untransmitted_proband, Transmitted_proband_parents, Untransmitted_proband_parents)
OR=numeric(); pvalue=numeric(); CI_lower=numeric(); CI_upper=numeric()

######## category 1; private variants: total allele in parents equal to 1   
Trans_variant=rare_syn_variants_new_sample %>% filter((Transmitted_proband_parents+Untransmitted_proband_parents)==1) %>%select(Transmitted_proband)%>%pull
Untrans_variant=rare_syn_variants_new_sample%>% filter((Transmitted_proband_parents+Untransmitted_proband_parents)==1)%>%select(Untransmitted_proband)%>%pull
fisher_test=fisher.test(matrix(c(sum(Trans_variant), num_proband, sum(Untrans_variant), num_proband), nrow=2))
OR[1]=fisher_test$estimate; pvalue[1]=fisher_test$p.value; CI_lower[1]=fisher_test$conf.int[1]; CI_upper[1]=fisher_test$conf.int[2]


######## category 2 1%<AF<5%
Trans_variant=rare_syn_variants_new_sample %>% filter(Gnomad_non_neuro_AF>0.01)%>%select(Transmitted_proband)%>%pull
Untrans_variant=rare_syn_variants_new_sample %>% filter(Gnomad_non_neuro_AF>0.01)%>%select(Untransmitted_proband)%>%pull
fisher_test=fisher.test(matrix(c(sum(Trans_variant), num_proband, sum(Untrans_variant), num_proband), nrow=2))
OR[2]=fisher_test$estimate; pvalue[2]=fisher_test$p.value; CI_lower[2]=fisher_test$conf.int[1]; CI_upper[2]=fisher_test$conf.int[2]


######## category 3 
Trans_variant=rare_syn_variants_new_sample %>% filter(Gnomad_non_neuro_AF>0.001 & Gnomad_non_neuro_AF<0.01)%>%select(Transmitted_proband)%>%pull
Untrans_variant=rare_syn_variants_new_sample %>% filter(Gnomad_non_neuro_AF>0.001 & Gnomad_non_neuro_AF<0.01)%>%select(Untransmitted_proband)%>%pull
fisher_test=fisher.test(matrix(c(sum(Trans_variant), num_proband, sum(Untrans_variant), num_proband), nrow=2))
OR[3]=fisher_test$estimate; pvalue[3]=fisher_test$p.value; CI_lower[3]=fisher_test$conf.int[1]; CI_upper[3]=fisher_test$conf.int[2]

######## category 4 
Trans_variant=rare_syn_variants_new_sample %>% filter(Gnomad_non_neuro_AF<0.001)%>%select(Transmitted_proband)%>%pull
Untrans_variant=rare_syn_variants_new_sample %>% filter(Gnomad_non_neuro_AF<0.001)%>%select(Untransmitted_proband)%>%pull
fisher_test=fisher.test(matrix(c(sum(Trans_variant), num_proband, sum(Untrans_variant), num_proband), nrow=2))
OR[4]=fisher_test$estimate; pvalue[4]=fisher_test$p.value; CI_lower[4]=fisher_test$conf.int[1]; CI_upper[4]=fisher_test$conf.int[2]

syn_varint_burden=tibble(OR=c(OR[2:4], OR[1]), var=c("(1): 1%<AF<5%", "(2): 0.1%<AF<1%", "(3): AF<0.1%", "(4): private"), lower=c(CI_lower[2:4], CI_lower[1]), upper=c(CI_upper[2:4], CI_upper[1]))

ggplot(syn_varint_burden, aes(x=var, y=OR, fill=var)) + # draw the OR plot 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  ylim(c(0,1))+
  ggtitle("Burden of syn variants")+
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.2,
                 position=position_dodge(.9))+
  theme(plot.title = element_text(hjust = 0.5))+ # center the title
  theme(legend.position="none")

```



#### filter `VQSLOD`

* `Gnomad_non_neuro_AF<0.05`
* `Consequence=="synonymous_variant"`
* after applying the above two filters, maximum of `VQSLOD` is 1.91, thus no need to use `cutoff of 3.75, but then brought it up to 5.13 ` suggested by Kyle. 


```{r, echo=F, warning=F, message=F, eval=T}
######### obtain syn missense variants in new sample data

rare_syn_variant_new_sample=ASC_new_sample %>% filter(Consequence=="synonymous_variant" & Gnomad_non_neuro_AF<0.05) %>% select(Variant, Gene, Transmitted_proband, Untransmitted_proband)  # find syn variants   in new sample data with Gnomad AF threshold 

transmitted_new_sample=numeric()
nontransmitted_new_sample=numeric()
for ( i in 1:length(Gene_Set_List))
 {
  Feature_Data=rare_syn_variant_new_sample%>%filter(Gene %in% Gene_Set_List[[i]])%>% drop_na()
  if (nrow(Feature_Data)>0)
  { 
    transmitted_new_sample[i]=Feature_Data%>%select(Transmitted_proband)%>%sum()
    nontransmitted_new_sample[i]=Feature_Data%>%select(Untransmitted_proband)%>%sum()
    
     ##############  run mirage-vs on every variant set, takes several minutes to run
#  Feature_Data_VS=cell_variant_set%>% filter(GENE_NAME %in% Gene_Set_List[[i]])%>% select(Transmitted_proband,Untransmitted_proband) %>% add_column(Group_Index=1)%>% drop_na()
#  Burden_MIRAGE_VS[i,1:2]=c(transmitted_cell_sample[i], nontransmitted_cell_sample[i])
#  pois.test=poisson.test(Burden_MIRAGE_VS[i,1:2], c(5869, 5869), r=1, alternative="greater")
# Burden_MIRAGE_VS[i,3:4]=c(pois.test$estimate,pois.test$p.value)
 
#  vs.mirage=mirage_vs(data.frame(Feature_Data_VS),n1=5869,n2=5869)
#  Burden_MIRAGE_VS[i,5:6]=c(vs.mirage$eta.est, vs.mirage$eta.pvalue)
  
  ##############
    
  }
  }  ### end of i

#write.csv(Burden_MIRAGE_VS, file="../output/CellDataASC/Variant_Set_missense_10Geneset.csv")

syn_new_sample=tibble(Gene=rep(Gene_Set,2), status=rep(c("transmitted", "nontransmitted"), each=10), variant_count=c(transmitted_new_sample, nontransmitted_new_sample))

g_new_sample=ggplot(syn_new_sample, aes(x = Gene, y = variant_count, fill = status)) +
  geom_col(position = position_dodge())+
  #geom_bar(stat="identity")+
  theme_classic()+
  ylim(c(0,40000))+
  ylab("Variant count")+xlab("Gene set")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=10))+
ggtitle("syn variants in new sample")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title
g_new_sample
############ combine two figures into one panel 
#figure=ggarrange( g_old_sample, g_new_sample, common.legend = TRUE, legend="right") 
#annotate_figure(figure,
#                top = text_grob("missense variants; AF<0.0001", color = "red", face = "bold", size = 14),
#                bottom = text_grob("Data source: \n ToothGrowth data set", color = "blue",
#                                   hjust = 1, x = 1, face = "italic", size = 10),
#                left = text_grob("Tooth length", color = "green", rot = 90),
#                right = "I'm done, thanks :-)!",
#                fig.lab = "Figure 1", fig.lab.face = "bold"
#)
```


* fisher exact test is used 
* lower, upper indicate 95% confidence interval 

```{r, echo=F, message=F, warning=F}
#################### print out summary statistics in a  table 
syn_new_sample_summary=tibble()
p.value=numeric(); lower_bound=numeric(); upper_bound=numeric()
num_proband=6026+1265 # male proband+female proband 
for (i in 1:length(transmitted_new_sample))
{
  fisher_test=fisher.test(matrix(c(transmitted_new_sample[i], num_proband, nontransmitted_new_sample[i], num_proband), nrow=2))
  p.value[i]=fisher_test$p.val
  lower_bound[i]=fisher_test$conf.int[1]
  upper_bound[i]=fisher_test$conf.int[2]
}
syn_new_sample_summary=tibble(gene=Gene_Set,transmitted=transmitted_new_sample, nontransmitted=nontransmitted_new_sample, pvalue=p.value, lower=lower_bound, upper=upper_bound)
kable(syn_new_sample_summary, caption = "")%>%
kable_styling() %>%
scroll_box(height = "500px")
```


#### filter QD and AS_SOR

* `Gnomad_non_neuro_AF<0.05`
* `Consequence=="synonymous_variant"`
* `QD>=3  & AS_SOR <=3 & AS_ReadPosRankSum>-8`, filter used for indel by Kyle 

```{r, echo=F, warning=F, message=F, eval=T}
######### obtain syn missense variants in new sample data

rare_syn_variant_new_sample=ASC_new_sample %>% filter(Consequence=="synonymous_variant" & Gnomad_non_neuro_AF<0.05 & QD>=3  & AS_SOR <=3 & AS_ReadPosRankSum>-8) %>% select(Variant, Gene, Transmitted_proband, Untransmitted_proband)  # find syn variants   in new sample data with Gnomad AF threshold 

transmitted_new_sample=numeric()
nontransmitted_new_sample=numeric()
for ( i in 1:length(Gene_Set_List))
 {
  Feature_Data=rare_syn_variant_new_sample%>%filter(Gene %in% Gene_Set_List[[i]])%>% drop_na()
  if (nrow(Feature_Data)>0)
  { 
    transmitted_new_sample[i]=Feature_Data%>%select(Transmitted_proband)%>%sum()
    nontransmitted_new_sample[i]=Feature_Data%>%select(Untransmitted_proband)%>%sum()
    
     ##############  run mirage-vs on every variant set, takes several minutes to run
#  Feature_Data_VS=cell_variant_set%>% filter(GENE_NAME %in% Gene_Set_List[[i]])%>% select(Transmitted_proband,Untransmitted_proband) %>% add_column(Group_Index=1)%>% drop_na()
#  Burden_MIRAGE_VS[i,1:2]=c(transmitted_cell_sample[i], nontransmitted_cell_sample[i])
#  pois.test=poisson.test(Burden_MIRAGE_VS[i,1:2], c(5869, 5869), r=1, alternative="greater")
# Burden_MIRAGE_VS[i,3:4]=c(pois.test$estimate,pois.test$p.value)
 
#  vs.mirage=mirage_vs(data.frame(Feature_Data_VS),n1=5869,n2=5869)
#  Burden_MIRAGE_VS[i,5:6]=c(vs.mirage$eta.est, vs.mirage$eta.pvalue)
  
  ##############
    
  }
  }  ### end of i

#write.csv(Burden_MIRAGE_VS, file="../output/CellDataASC/Variant_Set_missense_10Geneset.csv")

syn_new_sample=tibble(Gene=rep(Gene_Set,2), status=rep(c("transmitted", "nontransmitted"), each=10), variant_count=c(transmitted_new_sample, nontransmitted_new_sample))

g_new_sample=ggplot(syn_new_sample, aes(x = Gene, y = variant_count, fill = status)) +
  geom_col(position = position_dodge())+
  #geom_bar(stat="identity")+
  theme_classic()+
  ylim(c(0,40000))+
  ylab("Variant count")+xlab("Gene set")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=10))+
ggtitle("syn variants in new sample")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title
g_new_sample
############ combine two figures into one panel 
#figure=ggarrange( g_old_sample, g_new_sample, common.legend = TRUE, legend="right") 
#annotate_figure(figure,
#                top = text_grob("missense variants; AF<0.0001", color = "red", face = "bold", size = 14),
#                bottom = text_grob("Data source: \n ToothGrowth data set", color = "blue",
#                                   hjust = 1, x = 1, face = "italic", size = 10),
#                left = text_grob("Tooth length", color = "green", rot = 90),
#                right = "I'm done, thanks :-)!",
#                fig.lab = "Figure 1", fig.lab.face = "bold"
#)

```


* fisher exact test is used 
* lower, upper indicate 95% confidence interval 

```{r, echo=F, message=F, warning=F}
#################### print out summary statistics in a  table 
syn_new_sample_summary=tibble()
p.value=numeric(); lower_bound=numeric(); upper_bound=numeric()
num_proband=6026+1265 # male proband+female proband 
for (i in 1:length(transmitted_new_sample))
{
  fisher_test=fisher.test(matrix(c(transmitted_new_sample[i], num_proband, nontransmitted_new_sample[i], num_proband), nrow=2))
  p.value[i]=fisher_test$p.val
  lower_bound[i]=fisher_test$conf.int[1]
  upper_bound[i]=fisher_test$conf.int[2]
}
syn_new_sample_summary=tibble(gene=Gene_Set,transmitted=transmitted_new_sample, nontransmitted=nontransmitted_new_sample, pvalue=p.value, lower=lower_bound, upper=upper_bound)
kable(syn_new_sample_summary, caption = "")%>%
kable_styling() %>%
scroll_box(height = "500px")
```

#### filter QD and AS_SOR 


* `Gnomad_non_neuro_AF<0.05`
* `Consequence=="synonymous_variant"`
* `QD>=1  & AS_SOR <=3 & AS_ReadPosRankSum>-0.8`, filter used for SNP by Kyle


```{r, echo=F, message=F, warning=F}
rare_syn_variant_new_sample=ASC_new_sample %>% filter(Consequence=="synonymous_variant" & Gnomad_non_neuro_AF<0.05 & QD>=1  & AS_SOR <=3 & AS_ReadPosRankSum>-0.8) %>% select(Variant, Gene, Transmitted_proband, Untransmitted_proband)  # find syn variants   in new sample data with Gnomad AF threshold 


fisher_test=fisher.test(matrix(c(sum(rare_syn_variant_new_sample$Transmitted_proband), num_proband, sum(rare_syn_variant_new_sample$Untransmitted_proband), num_proband), nrow=2))
fisher_test_result_syn=tibble(OR=fisher_test$estimate, lower=fisher_test$conf.int[1], upper=fisher_test$conf.int[2], var="syn")
ggplot(fisher_test_result_syn, aes(x=var, y=OR, fill=var)) + # draw the OR plot 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.2,
                 position=position_dodge(.9)) 
```


```{r, echo=F, warning=F, message=F, eval=T}
######### obtain syn missense variants in new sample data

rare_syn_variant_new_sample=ASC_new_sample %>% filter(Consequence=="synonymous_variant" & Gnomad_non_neuro_AF<0.05 & QD>=1  & AS_SOR <=3 & AS_ReadPosRankSum>-0.8) %>% select(Variant, Gene, Transmitted_proband, Untransmitted_proband)  # find syn variants   in new sample data with Gnomad AF threshold 

transmitted_new_sample=numeric()
nontransmitted_new_sample=numeric()
for ( i in 1:length(Gene_Set_List))
 {
  Feature_Data=rare_syn_variant_new_sample%>%filter(Gene %in% Gene_Set_List[[i]])%>% drop_na()
  if (nrow(Feature_Data)>0)
  { 
    transmitted_new_sample[i]=Feature_Data%>%select(Transmitted_proband)%>%sum()
    nontransmitted_new_sample[i]=Feature_Data%>%select(Untransmitted_proband)%>%sum()
    
     ##############  run mirage-vs on every variant set, takes several minutes to run
#  Feature_Data_VS=cell_variant_set%>% filter(GENE_NAME %in% Gene_Set_List[[i]])%>% select(Transmitted_proband,Untransmitted_proband) %>% add_column(Group_Index=1)%>% drop_na()
#  Burden_MIRAGE_VS[i,1:2]=c(transmitted_cell_sample[i], nontransmitted_cell_sample[i])
#  pois.test=poisson.test(Burden_MIRAGE_VS[i,1:2], c(5869, 5869), r=1, alternative="greater")
# Burden_MIRAGE_VS[i,3:4]=c(pois.test$estimate,pois.test$p.value)
 
#  vs.mirage=mirage_vs(data.frame(Feature_Data_VS),n1=5869,n2=5869)
#  Burden_MIRAGE_VS[i,5:6]=c(vs.mirage$eta.est, vs.mirage$eta.pvalue)
  
  ##############
    
  }
  }  ### end of i

#write.csv(Burden_MIRAGE_VS, file="../output/CellDataASC/Variant_Set_missense_10Geneset.csv")

syn_new_sample=tibble(Gene=rep(Gene_Set,2), status=rep(c("transmitted", "nontransmitted"), each=10), variant_count=c(transmitted_new_sample, nontransmitted_new_sample))

g_new_sample=ggplot(syn_new_sample, aes(x = Gene, y = variant_count, fill = status)) +
  geom_col(position = position_dodge())+
  #geom_bar(stat="identity")+
  theme_classic()+
  ylim(c(0,30000))+
  ylab("Variant count")+xlab("Gene set")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=10))+
ggtitle("syn variants in new sample")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title
g_new_sample
############ combine two figures into one panel 
#figure=ggarrange( g_old_sample, g_new_sample, common.legend = TRUE, legend="right") 
#annotate_figure(figure,
#                top = text_grob("missense variants; AF<0.0001", color = "red", face = "bold", size = 14),
#                bottom = text_grob("Data source: \n ToothGrowth data set", color = "blue",
#                                   hjust = 1, x = 1, face = "italic", size = 10),
#                left = text_grob("Tooth length", color = "green", rot = 90),
#                right = "I'm done, thanks :-)!",
#                fig.lab = "Figure 1", fig.lab.face = "bold"
#)

```


* fisher exact test is used 
* lower, upper indicate 95% confidence interval 

```{r, echo=F, message=F, warning=F}
#################### print out summary statistics in a  table 
syn_new_sample_summary=tibble()
p.value=numeric(); lower_bound=numeric(); upper_bound=numeric()
num_proband=6026+1265 # male proband+female proband 
for (i in 1:length(transmitted_new_sample))
{
  fisher_test=fisher.test(matrix(c(transmitted_new_sample[i], num_proband, nontransmitted_new_sample[i], num_proband), nrow=2))
  p.value[i]=fisher_test$p.val
  lower_bound[i]=fisher_test$conf.int[1]
  upper_bound[i]=fisher_test$conf.int[2]
}
syn_new_sample_summary=tibble(gene=Gene_Set,transmitted=transmitted_new_sample, nontransmitted=nontransmitted_new_sample, pvalue=p.value, lower=lower_bound, upper=upper_bound)
kable(syn_new_sample_summary, caption = "")%>%
kable_styling() %>%
scroll_box(height = "500px")
```


## Old sample vs new sample


### Missense variants in different gene set


To compare old sample and new sample


*  look at missense variant with AF<1E-4 in 10 gene sets

* Old sample: missense: Non-LoF variants  

* new sample:  there are 2 labels related to "missense". one is `missense_variant` in column `Consequence` and the other is column `isMis`. All 100201 `missense_variant` are with `isMis=TRUE`, but not all 870738 `isMis==TRUE`  are `missense_variant`. Use `missense_variant` as missense variants  and the filter `QD>=1  & AS_SOR <=3 & AS_ReadPosRankSum>-0.8`

```{r, echo=F, eval=F}
sum(ASC_new_sample2%>% select(Consequence, isMis) %>% filter(Consequence=="missense_variant") %>%select(isMis)%>%pull() # how many TRUE isMis in misense_variant
ASC_new_sample2%>% select(Consequence, isMis) %>% filter(isMis=="TRUE") # how many TRUE isMis    
```



#### Variant counts comparison


```{r, echo=F, warning=F, message=F, cache=T}
######### obtain missense variants with AF<1-E4+10 gene set in old sample
LoF.def=c("stopgain", "frameshift substitution", "splicing", "stoploss")
############## to save burden and mirage-vs result 
Burden_MIRAGE_VS=matrix(nrow=length(Gene_Set_List), ncol=6)
rownames(Burden_MIRAGE_VS)=Gene_Set
colnames(Burden_MIRAGE_VS)=c("Transmitted", "Untransmitted", "Burden_OR", "Burden_Pval", "MIRAGE-VS_eta", "MIRAGE-VS_Pval")
##############
#old_variant_set=Anno.Data%>%filter(as.character(Polyphen2.HDIV.score)>0 & ExacAF<0.0001)%>%select(ID, Gene, No.case, No.contr) %>% drop_na() # missense variant with AF<1E-4 in old sample 
rare_missense_variant_old=Anno.Data%>%filter(Annotation %in% LoF.def==F & ExacAF<0.0001)%>%select(ID, Gene, No.case, No.contr) %>% drop_na() # missense variant with AF<1E-4 in old sample
transmitted_old_sample=numeric()
nontransmitted_old_sample=numeric()
for ( i in 1:length(Gene_Set_List))
 {  #i=1
  Feature_Data=rare_missense_variant_old%>%filter(Gene %in% Gene_Set_List[[i]])
  if (nrow(Feature_Data)>0)
  { 
    transmitted_old_sample[i]=Feature_Data%>%select(No.case)%>%sum()
    nontransmitted_old_sample[i]=Feature_Data%>%select(No.contr)%>%sum()
  }
  ##############  run mirage-vs on every variant set, takes several minutes to run
#  Feature_Data_VS=old_variant_set%>% filter(Gene %in% Gene_Set_List[[i]])%>% select(No.case,No.contr) %>% add_column(Group_Index=1)%>% drop_na()
#  Burden_MIRAGE_VS[i,1:2]=c(transmitted_old_sample[i], nontransmitted_old_sample[i])
#  pois.test=poisson.test(Burden_MIRAGE_VS[i,1:2], c(4315, 4315), r=1, alternative="greater")
# Burden_MIRAGE_VS[i,3:4]=c(pois.test$estimate,pois.test$p.value)
 
#  vs.mirage=mirage_vs(data.frame(Feature_Data_VS),n1=4315,n2=4315)
#  Burden_MIRAGE_VS[i,5:6]=c(vs.mirage$eta.est, vs.mirage$eta.pvalue)
  
  ##############
  }  # end of i

#write.csv(Burden_MIRAGE_VS, file="../output/OldSample/Variant_Set_missense_10Geneset.csv")


missense_old_sample=tibble(Gene=rep(Gene_Set,2), status=rep(c("transmitted", "nontransmitted"), each=10), variant_count=c(transmitted_old_sample, nontransmitted_old_sample))

g_old_sample=ggplot(missense_old_sample, aes(x = Gene, y = variant_count, fill = status)) +
  geom_col(position = "dodge")+
  theme_classic()+
  ylim(c(0,41000))+
  ylab("Variant count")+xlab("Gene set")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=10))+
ggtitle("old sample")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title


```





```{r, echo=F, warning=F, message=F, eval=T}
######### obtain missense variants with AF<1-E4+10 gene set in new sample data
############## to save burden and mirage-vs result 
Burden_MIRAGE_VS=matrix(nrow=length(Gene_Set_List), ncol=6)
rownames(Burden_MIRAGE_VS)=Gene_Set
colnames(Burden_MIRAGE_VS)=c("Transmitted", "Untransmitted", "Burden_OR", "Burden_Pval", "MIRAGE-VS_eta", "MIRAGE-VS_Pval")
##############

rare_missense_variant_new_sample=ASC_new_sample %>% filter(Consequence %in% c("missense_variant,NMD_transcript_variant","missense_variant", "missense_variant,splice_region_variant")==T )%>% filter(Gnomad_non_neuro_AF<0.0001 & QD>=1  & AS_SOR <=3 & AS_ReadPosRankSum>-0.8) %>% select(Variant, Gene, Transmitted_proband, Untransmitted_proband)  # find missense variant in new sample data with AF<1-10-4

transmitted_new_sample=numeric()
nontransmitted_new_sample=numeric()
for ( i in 1:length(Gene_Set_List))
 {
  Feature_Data=rare_missense_variant_new_sample%>%filter(Gene %in% Gene_Set_List[[i]])%>% drop_na()
  if (nrow(Feature_Data)>0)
  { 
    transmitted_new_sample[i]=Feature_Data%>%select(Transmitted_proband)%>%sum()
    nontransmitted_new_sample[i]=Feature_Data%>%select(Untransmitted_proband)%>%sum()
    
     ##############  run mirage-vs on every variant set, takes several minutes to run
#  Feature_Data_VS=cell_variant_set%>% filter(GENE_NAME %in% Gene_Set_List[[i]])%>% select(Transmitted_proband,Untransmitted_proband) %>% add_column(Group_Index=1)%>% drop_na()
#  Burden_MIRAGE_VS[i,1:2]=c(transmitted_cell_sample[i], nontransmitted_cell_sample[i])
#  pois.test=poisson.test(Burden_MIRAGE_VS[i,1:2], c(5869, 5869), r=1, alternative="greater")
# Burden_MIRAGE_VS[i,3:4]=c(pois.test$estimate,pois.test$p.value)
 
#  vs.mirage=mirage_vs(data.frame(Feature_Data_VS),n1=5869,n2=5869)
#  Burden_MIRAGE_VS[i,5:6]=c(vs.mirage$eta.est, vs.mirage$eta.pvalue)
  
  ##############
    
  }
  }  ### end of i

#write.csv(Burden_MIRAGE_VS, file="../output/CellDataASC/Variant_Set_missense_10Geneset.csv")

missense_new_sample=tibble(Gene=rep(Gene_Set,2), status=rep(c("transmitted", "nontransmitted"), each=10), variant_count=c(transmitted_new_sample, nontransmitted_new_sample))

g_new_sample=ggplot(missense_new_sample, aes(x = Gene, y = variant_count, fill = status)) +
  geom_col(position = position_dodge())+
  #geom_bar(stat="identity")+
  theme_classic()+
  ylim(c(0,41000))+
  ylab("Variant count")+xlab("Gene set")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=10))+
ggtitle("new sample")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title

figure=ggarrange( g_old_sample, g_new_sample, common.legend = TRUE, legend="right") 
annotate_figure(figure,
                top = text_grob("missense variants; AF<0.0001", color = "red", face = "bold", size = 14),
#                bottom = text_grob("Data source: \n ToothGrowth data set", color = "blue",
#                                   hjust = 1, x = 1, face = "italic", size = 10),
#                left = text_grob("Tooth length", color = "green", rot = 90),
#                right = "I'm done, thanks :-)!",
#                fig.lab = "Figure 1", fig.lab.face = "bold"
)
```


* `Gnomad_non_neuro_AF<0.05`
*  missense variants include `"missense_variant,NMD_transcript_variant","missense_variant", "missense_variant,splice_region_variant"`
* use filter `QD>=1  & AS_SOR <=3 & AS_ReadPosRankSum>-0.8`

```{r, echo=F, message=F, warning=F}
## burden for missense variants 
rare_missense_variant_new_sample=ASC_new_sample %>% filter(Consequence %in% c("missense_variant,NMD_transcript_variant","missense_variant", "missense_variant,splice_region_variant")==T )%>% filter(Gnomad_non_neuro_AF<0.05 & QD>=1  & AS_SOR <=3 & AS_ReadPosRankSum>-0.8) %>% select(Variant, Gene, Transmitted_proband, Untransmitted_proband)  # find missense variant in new sample data with AF filter 

fisher_test=fisher.test(matrix(c(sum(rare_missense_variant_new_sample$Transmitted_proband), num_proband, sum(rare_missense_variant_new_sample$Untransmitted_proband), num_proband), nrow=2))
fisher_test_result_mis=tibble(OR=fisher_test$estimate, lower=fisher_test$conf.int[1], upper=fisher_test$conf.int[2], var="mis")
ggplot(fisher_test_result_mis, aes(x=var, y=OR, fill=var)) + # draw the OR plot 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.2,
                 position=position_dodge(.9)) 

```

```{r, echo=F, message=F, warning=F}
# combine burden for all variants, syn and missense 
fisher_test_result=rbind(fisher_test_result_all, fisher_test_result_syn, fisher_test_result_mis)
ggplot(fisher_test_result, aes(x=var, y=OR, fill=var)) + # draw the OR plot 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.2,
                 position=position_dodge(.9)) 
```


#### $10^{-4}< AF<10^{-3}$


```{r, echo=F, warning=F, message=F, cache=T}
######### obtain missense variants with AF<1-E4+10 gene set in old sample
LoF.def=c("stopgain", "frameshift substitution", "splicing", "stoploss")
############## to save burden and mirage-vs result 
Burden_MIRAGE_VS=matrix(nrow=length(Gene_Set_List), ncol=6)
rownames(Burden_MIRAGE_VS)=Gene_Set
colnames(Burden_MIRAGE_VS)=c("Transmitted", "Untransmitted", "Burden_OR", "Burden_Pval", "MIRAGE-VS_eta", "MIRAGE-VS_Pval")
##############
#old_variant_set=Anno.Data%>%filter(as.character(Polyphen2.HDIV.score)>0 & ExacAF<0.0001)%>%select(ID, Gene, No.case, No.contr) %>% drop_na() # missense variant with AF<1E-4 in old sample 
rare_missense_variant_old=Anno.Data%>%filter(Annotation %in% LoF.def==F & ExacAF<0.001 & ExacAF>0.0001 )%>%select(ID, Gene, No.case, No.contr) %>% drop_na() # missense variant with AF<1E-4 in old sample
transmitted_old_sample=numeric()
nontransmitted_old_sample=numeric()
for ( i in 1:length(Gene_Set_List))
 {  #i=1
  Feature_Data=rare_missense_variant_old%>%filter(Gene %in% Gene_Set_List[[i]])
  if (nrow(Feature_Data)>0)
  { 
    transmitted_old_sample[i]=Feature_Data%>%select(No.case)%>%sum()
    nontransmitted_old_sample[i]=Feature_Data%>%select(No.contr)%>%sum()
  }
  ##############  run mirage-vs on every variant set, takes several minutes to run
#  Feature_Data_VS=old_variant_set%>% filter(Gene %in% Gene_Set_List[[i]])%>% select(No.case,No.contr) %>% add_column(Group_Index=1)%>% drop_na()
#  Burden_MIRAGE_VS[i,1:2]=c(transmitted_old_sample[i], nontransmitted_old_sample[i])
#  pois.test=poisson.test(Burden_MIRAGE_VS[i,1:2], c(4315, 4315), r=1, alternative="greater")
# Burden_MIRAGE_VS[i,3:4]=c(pois.test$estimate,pois.test$p.value)
 
#  vs.mirage=mirage_vs(data.frame(Feature_Data_VS),n1=4315,n2=4315)
#  Burden_MIRAGE_VS[i,5:6]=c(vs.mirage$eta.est, vs.mirage$eta.pvalue)
  
  ##############
  }  # end of i

#write.csv(Burden_MIRAGE_VS, file="../output/OldSample/Variant_Set_missense_10Geneset.csv")


missense_old_sample=tibble(Gene=rep(Gene_Set,2), status=rep(c("transmitted", "nontransmitted"), each=10), variant_count=c(transmitted_old_sample, nontransmitted_old_sample))

g_old_sample=ggplot(missense_old_sample, aes(x = Gene, y = variant_count, fill = status)) +
  geom_col(position = "dodge")+
  theme_classic()+
  ylim(c(0,35000))+
  ylab("Variant count")+xlab("Gene set")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=10))+
ggtitle("old sample")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title


```


```{r, echo=F, warning=F, message=F, eval=T}
######### obtain missense variants with AF<1-E4+10 gene set in new sample data
############## to save burden and mirage-vs result 
Burden_MIRAGE_VS=matrix(nrow=length(Gene_Set_List), ncol=6)
rownames(Burden_MIRAGE_VS)=Gene_Set
colnames(Burden_MIRAGE_VS)=c("Transmitted", "Untransmitted", "Burden_OR", "Burden_Pval", "MIRAGE-VS_eta", "MIRAGE-VS_Pval")
##############

rare_missense_variant_new_sample=ASC_new_sample %>% filter(Consequence %in% c("missense_variant,NMD_transcript_variant","missense_variant", "missense_variant,splice_region_variant")==T )%>% filter(Gnomad_non_neuro_AF<0.001 & Gnomad_non_neuro_AF>0.0001 & QD>=1  & AS_SOR <=3 & AS_ReadPosRankSum>-0.8) %>% select(Variant, Gene, Transmitted_proband, Untransmitted_proband)  # find missense variant in new sample data with AF<1-10-4

transmitted_new_sample=numeric()
nontransmitted_new_sample=numeric()
for ( i in 1:length(Gene_Set_List))
 {
  Feature_Data=rare_missense_variant_new_sample%>%filter(Gene %in% Gene_Set_List[[i]])%>% drop_na()
  if (nrow(Feature_Data)>0)
  { 
    transmitted_new_sample[i]=Feature_Data%>%select(Transmitted_proband)%>%sum()
    nontransmitted_new_sample[i]=Feature_Data%>%select(Untransmitted_proband)%>%sum()
    
     ##############  run mirage-vs on every variant set, takes several minutes to run
#  Feature_Data_VS=cell_variant_set%>% filter(GENE_NAME %in% Gene_Set_List[[i]])%>% select(Transmitted_proband,Untransmitted_proband) %>% add_column(Group_Index=1)%>% drop_na()
#  Burden_MIRAGE_VS[i,1:2]=c(transmitted_cell_sample[i], nontransmitted_cell_sample[i])
#  pois.test=poisson.test(Burden_MIRAGE_VS[i,1:2], c(5869, 5869), r=1, alternative="greater")
# Burden_MIRAGE_VS[i,3:4]=c(pois.test$estimate,pois.test$p.value)
 
#  vs.mirage=mirage_vs(data.frame(Feature_Data_VS),n1=5869,n2=5869)
#  Burden_MIRAGE_VS[i,5:6]=c(vs.mirage$eta.est, vs.mirage$eta.pvalue)
  
  ##############
    
  }
  }  ### end of i

#write.csv(Burden_MIRAGE_VS, file="../output/CellDataASC/Variant_Set_missense_10Geneset.csv")

missense_new_sample=tibble(Gene=rep(Gene_Set,2), status=rep(c("transmitted", "nontransmitted"), each=10), variant_count=c(transmitted_new_sample, nontransmitted_new_sample))

g_new_sample=ggplot(missense_new_sample, aes(x = Gene, y = variant_count, fill = status)) +
  geom_col(position = position_dodge())+
  #geom_bar(stat="identity")+
  theme_classic()+
  ylim(c(0,35000))+
  ylab("Variant count")+xlab("Gene set")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=10))+
ggtitle("new sample")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title

figure=ggarrange(g_old_sample, g_new_sample, common.legend = TRUE, legend="right") 
annotate_figure(figure,
                top = text_grob("missense variants; 0.0001<AF<0.001", color = "red", face = "bold", size = 14),
#                bottom = text_grob("Data source: \n ToothGrowth data set", color = "blue",
#                                   hjust = 1, x = 1, face = "italic", size = 10),
#                left = text_grob("Tooth length", color = "green", rot = 90),
#                right = "I'm done, thanks :-)!",
#                fig.lab = "Figure 1", fig.lab.face = "bold"
)
```


#### Missense: isMis==TRUE


This definition of missense variant is very likely not correct! 


```{r, echo=F, warning=F, message=F, eval=F}
######### obtain missense variants with AF<1-E4+10 gene set in new sample data
############## to save burden and mirage-vs result 
Burden_MIRAGE_VS=matrix(nrow=length(Gene_Set_List), ncol=6)
rownames(Burden_MIRAGE_VS)=Gene_Set
colnames(Burden_MIRAGE_VS)=c("Transmitted", "Untransmitted", "Burden_OR", "Burden_Pval", "MIRAGE-VS_eta", "MIRAGE-VS_Pval")
##############

rare_missense_variant_new_sample=ASC_new_sample %>% filter(isMis==TRUE)%>% filter(Gnomad_non_neuro_AF<0.0001) %>% select(Variant, Gene, Transmitted_proband, Untransmitted_proband)  # missense variant in new sample data with AF<1-10-4, define missense as isMis=TRUE

transmitted_new_sample=numeric()
nontransmitted_new_sample=numeric()
for ( i in 1:length(Gene_Set_List))
 {
  Feature_Data=rare_missense_variant_new_sample%>%filter(Gene %in% Gene_Set_List[[i]])%>% drop_na()
  if (nrow(Feature_Data)>0)
  { 
    transmitted_new_sample[i]=Feature_Data%>%select(Transmitted_proband)%>%sum()
    nontransmitted_new_sample[i]=Feature_Data%>%select(Untransmitted_proband)%>%sum()
    
     ##############  run mirage-vs on every variant set, takes several minutes to run
#  Feature_Data_VS=cell_variant_set%>% filter(GENE_NAME %in% Gene_Set_List[[i]])%>% select(Transmitted_proband,Untransmitted_proband) %>% add_column(Group_Index=1)%>% drop_na()
#  Burden_MIRAGE_VS[i,1:2]=c(transmitted_cell_sample[i], nontransmitted_cell_sample[i])
#  pois.test=poisson.test(Burden_MIRAGE_VS[i,1:2], c(5869, 5869), r=1, alternative="greater")
# Burden_MIRAGE_VS[i,3:4]=c(pois.test$estimate,pois.test$p.value)
 
#  vs.mirage=mirage_vs(data.frame(Feature_Data_VS),n1=5869,n2=5869)
#  Burden_MIRAGE_VS[i,5:6]=c(vs.mirage$eta.est, vs.mirage$eta.pvalue)
  
  ##############
    
  }
  }  ### end of i

#write.csv(Burden_MIRAGE_VS, file="../output/CellDataASC/Variant_Set_missense_10Geneset.csv")

missense_new_sample=tibble(Gene=rep(Gene_Set,2), status=rep(c("transmitted", "nontransmitted"), each=10), variant_count=c(transmitted_new_sample, nontransmitted_new_sample))

g_new_sample=ggplot(missense_new_sample, aes(x = Gene, y = variant_count, fill = status)) +
  geom_col(position = position_dodge())+
  #geom_bar(stat="identity")+
  theme_classic()+
  ylim(c(0,2300000))+
  ylab("Variant count")+xlab("Gene set")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=10))+
ggtitle("new sample")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title

#####################  this is from old sample 
g_old_sample=ggplot(missense_old_sample, aes(x = Gene, y = variant_count, fill = status)) +
  geom_col(position = "dodge")+
  theme_classic()+
  ylim(c(0,2300000))+
  ylab("Variant count")+xlab("Gene set")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=10))+
ggtitle("old sample")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title
#####################

ggarrange( g_old_sample, g_new_sample, common.legend = TRUE, legend="right") 
```


###  Missense variants in individual constraint gene 

consider missense variants with AF<10-E4 in top 10 genes from old sample and new sample and see how they compare 

```{r, echo=F}
Top_10_gene=c("CYFIP1", "EP400", "FBN1", "SRCAP", "DYNC1H1", "ABCA2", "DOCK4", "CACNA1D", "CHD8", "TRIP12")
### display  rare missense variants in individual top genes  
######################################################
## Gene_Set[7]: constraint gene 
# find rare missense variants in old sample and new sample 
variant_gene_old_sample=rare_missense_variant_old %>% filter(Gene %in% Gene_Set_List[[7]])%>% rename(GENE_NAME=Gene) 
variant_gene_new_sample=rare_missense_variant_new_sample %>% filter(Gene %in% Gene_Set_List[[7]])
variant_gene_new_sample=variant_gene_new_sample %>%rename(GENE_NAME=Gene) # rename a column for merge 
variant_gene_old_new_sample_combine=full_join(variant_gene_old_sample, variant_gene_new_sample, by="GENE_NAME") %>% select(GENE_NAME, No.case, No.contr, Transmitted_proband, Untransmitted_proband) # merge two data by common GENE_NAME 

####### fill NA with 0
variant_gene_old_new_sample_combine$No.case=variant_gene_old_new_sample_combine$No.case%>%replace_na(0)
variant_gene_old_new_sample_combine$No.contr=variant_gene_old_new_sample_combine$No.contr%>%replace_na(0)
variant_gene_old_new_sample_combine$Transmitted_proband=variant_gene_old_new_sample_combine$Transmitted_proband%>%replace_na(0)
variant_gene_old_new_sample_combine$Untransmitted_proband=variant_gene_old_new_sample_combine$Untransmitted_proband%>%replace_na(0)
########### count variants for every gene 
#unique_gene=unique(variant_gene_old_new_sample_combine$GENE_NAME)[1:50]
unique_gene=Top_10_gene
unique_gene_case=numeric()
unique_gene_control=numeric()
unique_gene_transmitted_proband=numeric()
unique_gene_untransmitted_proband=numeric()
for (i in 1:length(unique_gene))
{
 unique_gene_case[i]=variant_gene_old_new_sample_combine %>% filter(GENE_NAME %in% unique_gene[i]) %>% select(No.case)%>%sum()
 unique_gene_control[i]=variant_gene_old_new_sample_combine %>% filter(GENE_NAME %in% unique_gene[i]) %>% select(No.contr)%>%sum()
unique_gene_transmitted_proband[i]=variant_gene_old_new_sample_combine %>% filter(GENE_NAME %in% unique_gene[i]) %>% select(Transmitted_proband)%>%sum()  
unique_gene_untransmitted_proband[i]=variant_gene_old_new_sample_combine %>% filter(GENE_NAME %in% unique_gene[i]) %>% select(Untransmitted_proband)%>%sum()
}
unique_gene_variant_count=tibble(Gene=unique_gene, No.case=unique_gene_case, No.control=unique_gene_control, Transmitted_proband=unique_gene_transmitted_proband, Untransmitted_proband=unique_gene_untransmitted_proband)
###########################
transmitted_variant=tibble(Gene=c(as.character(unique_gene), as.character(unique_gene)), count=c(unique_gene_variant_count$No.case, unique_gene_variant_count$Transmitted_proband), sample=c(rep("old", nrow(unique_gene_variant_count)), rep("new", nrow(unique_gene_variant_count))))


g1=ggplot(transmitted_variant, aes(x=Gene, y=count, shape=sample, color=sample))+ 
 geom_point()+
  ylab("Transmitted variant counts")+
  #theme_classic()+
  theme(axis.text.x = element_text(angle = 60, vjust = 0.5, size=7))
g1

untransmitted_variant=tibble(Gene=c(as.character(unique_gene), as.character(unique_gene)), count=c(unique_gene_variant_count$No.control, unique_gene_variant_count$Untransmitted_proband), sample=c(rep("old", nrow(unique_gene_variant_count)), rep("new", nrow(unique_gene_variant_count))))


g2=ggplot(untransmitted_variant, aes(x=Gene, y=count, shape=sample, color=sample))+ 
 geom_point()+
  ylab("Untransmitted variant counts")+
  #theme_classic()+
  theme(axis.text.x = element_text(angle = 60, vjust = 0.5, size=7))
g2
```


## Gene-CYFIP1


* AF<0.05 by setting `Gnomad_non_neuro_AF<0.05`
* exclude synonymous variants by setting `pLI!=synonymous_variant`

```{r, echo=F}
CYFIP1=ASC_new_sample%>%filter(Gene=="CYFIP1" & pLI!="synonymous_variant"& Gnomad_non_neuro_AF<0.05) %>% select(Variant, Chrom, Gene, Consequence,Gnomad_non_neuro_AF, Transmitted_proband, Untransmitted_proband  )

kable(CYFIP1, caption = "Variant (AF<0.05) in gene CYFIP1")%>%
kable_styling() %>%
scroll_box(height = "500px")
```

In old sample, there is one variant having 8/0 at '15:22963816-22963816_A_G' which doesn’t show up in new sample.


* if no any constraint 
* all variants don't go beyond `2295xxxx`.

```{r, echo=F}
CYFIP1=ASC_new_sample%>%filter(Gene=="CYFIP1") %>% select(Variant, Chrom, Gene, Consequence,Gnomad_non_neuro_AF, Transmitted_proband, Untransmitted_proband  )

kable(CYFIP1, caption = "All variants  in gene CYFIP1")%>%
kable_styling() %>%
scroll_box(height = "500px")
```



### LoF variants in different gene set 



To compare old sample and new data, look at LoF +AF<1E-4+10 gene set and how they compare with each other. 

Old sample: LoF: the union of  "stopgain", "frameshift substitution", "splicing", "stoploss" 

new sample: LoF: 


```{r, echo=F, warning=F, message=F, eval=F}
######### obtain LOF variants with AF<1-E4+10 gene set in old sample
LoF.def=c("stopgain", "frameshift substitution", "splicing", "stoploss")
old_variant_set=Anno.Data%>%filter(Annotation %in% LoF.def & ExacAF<0.0001)%>%select(ID, Gene, No.case, No.contr) %>% drop_na() # LoF variant with AF<1E-4 in old sample
transmitted_old_sample=numeric()
nontransmitted_old_sample=numeric()
for ( i in 1:10)
 {
  Feature_Data=old_variant_set%>%filter(Gene %in% Gene_Set_List[[i]])
  if (nrow(Feature_Data)>0)
  { 
    transmitted_old_sample[i]=Feature_Data%>%select(No.case)%>%sum()
    nontransmitted_old_sample[i]=Feature_Data%>%select(No.contr)%>%sum()
  }
  }

LoF_old_sample=tibble(Gene=rep(Gene_Set,2), status=rep(c("transmitted", "nontransmitted"), each=10), variant_count=c(transmitted_old_sample, nontransmitted_old_sample))

g_old_sample=ggplot(LoF_old_sample, aes(x = Gene, y = variant_count, fill = status)) +
  geom_col(position = "dodge")+
  theme_classic()+
  ylim(c(0,1000))+
  ylab("Variant count")+xlab("Gene set")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=10))+
ggtitle("old sample")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title


```

```{r, echo=F, warning=F, message=F, eval=F}
######### obtain LOF variants with AF<1-E4+10 gene set in new data
cell_variant_set=cell_trans_data_AF %>% filter(pLI>=0.5) %>% filter(PTV==T)%>% filter(ExacAF<0.0001) %>% select(Variant, GENE_NAME, Transmitted_proband, Untransmitted_proband)  # LoF variant in cell data with AF<1-10-4
transmitted_cell_sample=numeric()
nontransmitted_cell_sample=numeric()
for ( i in 1:10)
 {
  Feature_Data=cell_variant_set%>%filter(GENE_NAME %in% Gene_Set_List[[i]])
  if (nrow(Feature_Data)>0)
  { 
    transmitted_cell_sample[i]=Feature_Data%>%select(Transmitted_proband)%>%sum()
    nontransmitted_cell_sample[i]=Feature_Data%>%select(Untransmitted_proband)%>%sum()
  }
  }

LoF_cell_sample=tibble(Gene=rep(Gene_Set,2), status=rep(c("transmitted", "nontransmitted"), each=10), variant_count=c(transmitted_cell_sample, nontransmitted_cell_sample))

g_cell_sample=ggplot(LoF_cell_sample, aes(x = Gene, y = variant_count, fill = status)) +
  geom_col(position = position_dodge())+
  #geom_bar(stat="identity")+
  theme_classic()+
  ylim(c(0,1000))+
  ylab("Variant count")+xlab("Gene set")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=10))+
ggtitle("cell sample")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title

ggarrange( g_old_sample, g_cell_sample, common.legend = TRUE, legend="right") 
```

