---
title: "Variant set (VS) analysis"
output: html_document
---


```{r, echo=F, warning=F, message=F}
rm(list=ls())
set.seed(123)
library(knitr)
library(RSQLite)
library(dplyr)
library(knitr)
library(kableExtra)
library(RColorBrewer)
library(gplots)
library(tidyverse)
library(gridExtra)
library(ggpubr)
#library("devtools")
#install.packages("Rtools")
#install_github('xinhe-lab/mirage')
library(mirage)
```



```{r burdern analysis-gene level, results='hide', echo=F, warning=F, message=F}
# read into gene set 
GeneDB=src_sqlite(path="C:\\Shengtong\\Research\\rare-var\\gene.list.db", create=F)
#GeneDB=src_sqlite(path="../../gene.list.db", create=F)
gene_cate1=data.frame(collect(tbl(GeneDB, "SFARI_HighConf")))
gene_cate2=data.frame(collect(tbl(GeneDB, "SFARI_StrongCand")))
gene_cate3=data.frame(collect(tbl(GeneDB, "SFARI_cate3_gene")))
gene_cate4=data.frame(collect(tbl(GeneDB, "SFARI_cate4_gene")))
gene_cate5=data.frame(collect(tbl(GeneDB, "SFARI_cate5_gene")))
gene_cate6=data.frame(collect(tbl(GeneDB, "SFARI_cate6_gene")))
gene_cateS=data.frame(collect(tbl(GeneDB, "SFARI_cateS_gene")))
TADAGene=data.frame(collect(tbl(GeneDB, "TADAGenelist")))
Qlessthan5percentgene=TADAGene$TadaName[TADAGene$qvalue.combined<0.05]
Qlessthan20percentgene=TADAGene$TadaName[TADAGene$qvalue.combined<0.2]
Qlessthan30percentgene=TADAGene$TadaName[TADAGene$qvalue.combined<0.3]
Qlessthan40percentgene=TADAGene$TadaName[TADAGene$qvalue.combined<0.4]
Qlessthan50percentgene=TADAGene$TadaName[TADAGene$qvalue.combined<0.5]
Qlargerthan90percentgene=TADAGene$TadaName[TADAGene$qvalue.combined>0.9]
###### ID gene ######
ID_Gene=as.character(data.frame(collect(tbl(GeneDB, "Pinto14AJHG_IDgene")))$GeneID)
#### high confidence gene 
High_Confidence_Gene=union(union(gene_cate1$GeneID, gene_cate2$GeneID),Qlessthan5percentgene)
#### moderate confidence gene 
Moderate_Confidence_Gene=setdiff(union(union(gene_cate3$GeneID, gene_cateS$GeneID), Qlessthan30percentgene),Qlessthan5percentgene)
#### psd gene 
purcell.genelist=data.frame(collect(tbl(GeneDB, "Purcell2014_genelist"))) ## PSD gene, SCZdenovo gene 
PSD_Gene=purcell.genelist$Gene_symbol[purcell.genelist$PSD=="Y"]
#### FMRP gene 
FMRP_Gene=purcell.genelist$Gene_symbol[purcell.genelist$FMRP.target=="Y"]
#### Autism KB gene 
ASD_KB_Gene=as.character(data.frame(collect(tbl(GeneDB, "AutismKB_gene")))$GeneID)
#### constraint gene 
Constraint_Gene=data.frame(collect(tbl(GeneDB, "Samocha_2014NG_constraintgene")))$gene
### RVIS gene 
RVIS.Allgene=data.frame(collect(tbl(GeneDB, "RVIS_gene")))
RVIS_Gene=RVIS.Allgene$GeneID[RVIS.Allgene$RVIS.percentile<5] # top 5% gene
#### Haplo insufficient gene 
Haploinsufficient_Gene=as.character(data.frame(collect(tbl(GeneDB, "Petrovski_plosgen_haploinsuff_gene")))$GeneID)
#### SCZ data 
SCZ_Gene=purcell.genelist$Gene_symbol
Gene_Set=c("ID gene","High", "Mod", "PSD", "FMRP", "AutismKB", "constraint gene", "RVIS", "Haploinsuff", "SCZ gene")
Gene_Set_List=list()
Gene_Set_List[[1]]=ID_Gene
Gene_Set_List[[2]]=High_Confidence_Gene
Gene_Set_List[[3]]=Moderate_Confidence_Gene
Gene_Set_List[[4]]=PSD_Gene
Gene_Set_List[[5]]=FMRP_Gene
Gene_Set_List[[6]]=ASD_KB_Gene
Gene_Set_List[[7]]=Constraint_Gene
Gene_Set_List[[8]]=RVIS_Gene
Gene_Set_List[[9]]=Haploinsufficient_Gene
Gene_Set_List[[10]]=SCZ_Gene
```



## Old sample vs new sample

```{r, echo=F, cache=T, warning=F, message=F}
### old sample
########## read into data
All.Anno.Data=as_tibble(read.table("../../AnnotatedTrans.txt", header=T))
N1=4315; N0=4315
All.Anno.Data[All.Anno.Data =="."] <- NA
All.Anno.Data$ExacAF[is.na(All.Anno.Data$ExacAF)]=0 # set AF of NA to zero 
Anno.Data=All.Anno.Data[which(All.Anno.Data$ExacAF<0.05 & All.Anno.Data$Annotation!="synonymous SNV"),] # use AF cutoff and exclude synonymous SNV
Anno.Data=All.Anno.Data[which(All.Anno.Data$Annotation!="synonymous SNV"),] # use AF cutoff and exclude synonumous SNV
```


```{r, echo=F, cache=T, warning=F, message=F}
### read into new sample 
ASC_new_sample1=as_tibble(read.table("C:\\Shengtong\\Research\\rare-var\\ASC\\ASC_B15_B16_raw_TDT_by_parent_2021-04-19.txt", header=T, fill=T))
ASC_new_sample2=as_tibble(read.table("C:\\Shengtong\\Research\\rare-var\\ASC\\ASC_v17_raw_TDT_by_parent_2021-05-10.txt", header=T, fill=T))
```


### Missense variants in different gene set


To compare old sample and new sample


*  look at missense variant with AF<1E-4 in 10 gene sets

* Old sample: missense: Non-LoF variants  

* new sample:  there are 2 labels related to "missense". one is `missense_variant` in column `Consequence` and the other is column `isMis`. All 100201 `missense_variant` are with `isMis=TRUE`, but not all 870738 `isMis==TRUE`  are `missense_variant`. Use `missense_variant` as missense variants 

```{r, echo=F, eval=F}
sum(ASC_new_sample2%>% select(Consequence, isMis) %>% filter(Consequence=="missense_variant") %>%select(isMis)%>%pull() # how many TRUE isMis in misense_variant
ASC_new_sample2%>% select(Consequence, isMis) %>% filter(isMis=="TRUE") # how many TRUE isMis    
```



#### Variant counts comparison


```{r, echo=F, warning=F, message=F, cache=T}
######### obtain missense variants with AF<1-E4+10 gene set in old sample
LoF.def=c("stopgain", "frameshift substitution", "splicing", "stoploss")
############## to save burden and mirage-vs result 
Burden_MIRAGE_VS=matrix(nrow=length(Gene_Set_List), ncol=6)
rownames(Burden_MIRAGE_VS)=Gene_Set
colnames(Burden_MIRAGE_VS)=c("Transmitted", "Untransmitted", "Burden_OR", "Burden_Pval", "MIRAGE-VS_eta", "MIRAGE-VS_Pval")
##############
#old_variant_set=Anno.Data%>%filter(as.character(Polyphen2.HDIV.score)>0 & ExacAF<0.0001)%>%select(ID, Gene, No.case, No.contr) %>% drop_na() # missense variant with AF<1E-4 in old sample 
rare_missense_variant_old=Anno.Data%>%filter(Annotation %in% LoF.def==F & ExacAF<0.0001)%>%select(ID, Gene, No.case, No.contr) %>% drop_na() # missense variant with AF<1E-4 in old sample
transmitted_old_sample=numeric()
nontransmitted_old_sample=numeric()
for ( i in 1:length(Gene_Set_List))
 {  #i=1
  Feature_Data=rare_missense_variant_old%>%filter(Gene %in% Gene_Set_List[[i]])
  if (nrow(Feature_Data)>0)
  { 
    transmitted_old_sample[i]=Feature_Data%>%select(No.case)%>%sum()
    nontransmitted_old_sample[i]=Feature_Data%>%select(No.contr)%>%sum()
  }
  ##############  run mirage-vs on every variant set, takes several minutes to run
#  Feature_Data_VS=old_variant_set%>% filter(Gene %in% Gene_Set_List[[i]])%>% select(No.case,No.contr) %>% add_column(Group_Index=1)%>% drop_na()
#  Burden_MIRAGE_VS[i,1:2]=c(transmitted_old_sample[i], nontransmitted_old_sample[i])
#  pois.test=poisson.test(Burden_MIRAGE_VS[i,1:2], c(4315, 4315), r=1, alternative="greater")
# Burden_MIRAGE_VS[i,3:4]=c(pois.test$estimate,pois.test$p.value)
 
#  vs.mirage=mirage_vs(data.frame(Feature_Data_VS),n1=4315,n2=4315)
#  Burden_MIRAGE_VS[i,5:6]=c(vs.mirage$eta.est, vs.mirage$eta.pvalue)
  
  ##############
  }  # end of i

#write.csv(Burden_MIRAGE_VS, file="../output/OldSample/Variant_Set_missense_10Geneset.csv")


missense_old_sample=tibble(Gene=rep(Gene_Set,2), status=rep(c("transmitted", "nontransmitted"), each=10), variant_count=c(transmitted_old_sample, nontransmitted_old_sample))

g_old_sample=ggplot(missense_old_sample, aes(x = Gene, y = variant_count, fill = status)) +
  geom_col(position = "dodge")+
  theme_classic()+
  ylim(c(0,50000))+
  ylab("Variant count")+xlab("Gene set")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=10))+
ggtitle("old sample")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title


```


```{r, echo=F, eval=T}
## identify transmitted and non-transmitted variants 
#transmitted_variant=ASC_new_sample2%>%select(starts_with(c("t", "u")), "Gnomad_non_neuro_AF")
#kable(transmitted_variant[1:10,])%>%
#kable_styling() %>%
#scroll_box(height = "500px")
#write.csv(transmitted_variant, file="C:\\Shengtong\\Research\\rare-var\\ASC\\ASC_v17_raw_TDT_by_parent_2021-05-10_transmitted_variant.csv")
##### count transmitted variants and untransmitted in proband by summing over male, female, dad, mom and ind 
Transmitted_proband=ASC_new_sample2%>% select(starts_with("t_proband"))%>%  mutate(Transmitted_proband = select(., t_proband_male_dad:t_proband_female_ind) %>% rowSums())%>% select(Transmitted_proband) 
Untransmitted_proband=ASC_new_sample2%>% select(starts_with("u_proband"))%>%  mutate(Untransmitted_proband = select(., u_proband_male_dad:u_proband_female_ind) %>% rowSums())%>% select(Untransmitted_proband) 

ASC_new_sample=ASC_new_sample2%>%mutate(Transmitted_proband=Transmitted_proband$Transmitted_proband, Untransmitted_proband=Untransmitted_proband$Untransmitted_proband)

```



```{r, echo=F, warning=F, message=F, eval=T}
######### obtain missense variants with AF<1-E4+10 gene set in new sample data
############## to save burden and mirage-vs result 
Burden_MIRAGE_VS=matrix(nrow=length(Gene_Set_List), ncol=6)
rownames(Burden_MIRAGE_VS)=Gene_Set
colnames(Burden_MIRAGE_VS)=c("Transmitted", "Untransmitted", "Burden_OR", "Burden_Pval", "MIRAGE-VS_eta", "MIRAGE-VS_Pval")
##############

rare_missense_variant_new_sample=ASC_new_sample %>% filter(Consequence=="missense_variant")%>% filter(Gnomad_non_neuro_AF<0.0001) %>% select(Variant, Gene, Transmitted_proband, Untransmitted_proband)  # missense variant in new sample data with AF<1-10-4

transmitted_new_sample=numeric()
nontransmitted_new_sample=numeric()
for ( i in 1:length(Gene_Set_List))
 {
  Feature_Data=rare_missense_variant_new_sample%>%filter(Gene %in% Gene_Set_List[[i]])%>% drop_na()
  if (nrow(Feature_Data)>0)
  { 
    transmitted_new_sample[i]=Feature_Data%>%select(Transmitted_proband)%>%sum()
    nontransmitted_new_sample[i]=Feature_Data%>%select(Untransmitted_proband)%>%sum()
    
     ##############  run mirage-vs on every variant set, takes several minutes to run
#  Feature_Data_VS=cell_variant_set%>% filter(GENE_NAME %in% Gene_Set_List[[i]])%>% select(Transmitted_proband,Untransmitted_proband) %>% add_column(Group_Index=1)%>% drop_na()
#  Burden_MIRAGE_VS[i,1:2]=c(transmitted_cell_sample[i], nontransmitted_cell_sample[i])
#  pois.test=poisson.test(Burden_MIRAGE_VS[i,1:2], c(5869, 5869), r=1, alternative="greater")
# Burden_MIRAGE_VS[i,3:4]=c(pois.test$estimate,pois.test$p.value)
 
#  vs.mirage=mirage_vs(data.frame(Feature_Data_VS),n1=5869,n2=5869)
#  Burden_MIRAGE_VS[i,5:6]=c(vs.mirage$eta.est, vs.mirage$eta.pvalue)
  
  ##############
    
  }
  }  ### end of i

#write.csv(Burden_MIRAGE_VS, file="../output/CellDataASC/Variant_Set_missense_10Geneset.csv")

missense_new_sample=tibble(Gene=rep(Gene_Set,2), status=rep(c("transmitted", "nontransmitted"), each=10), variant_count=c(transmitted_new_sample, nontransmitted_new_sample))

g_new_sample=ggplot(missense_new_sample, aes(x = Gene, y = variant_count, fill = status)) +
  geom_col(position = position_dodge())+
  #geom_bar(stat="identity")+
  theme_classic()+
  ylim(c(0,50000))+
  ylab("Variant count")+xlab("Gene set")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=10))+
ggtitle("new sample")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title

ggarrange( g_old_sample, g_new_sample, common.legend = TRUE, legend="right") 
```