---
title: "Cell data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r, echo=F}
rm(list=ls())
set.seed(123)
library(tidyverse)
library(knitr)
```

This is the transmitted data from [cell paper](https://www.sciencedirect.com/science/article/abs/pii/S0092867419313984)

```{r, echo=F, cache=T}
part1=as_tibble(read.table("../../AutismDatafromASC_Cellpaper/ASC_v15_TDT_GQPL25.hAB30.results.2018-04-01.part1.tsv.gz",sep = '\t', header = TRUE, fill = TRUE))
part2=as_tibble(read.table("../../AutismDatafromASC_Cellpaper/ASC_v15_TDT_GQPL25.hAB30.results.2018-04-01.part2.tsv.gz",sep = '\t', header = TRUE, fill = TRUE))
cell_trans_data=full_join(part1, part2)
```

```{r, echo=T}
cell_trans_data
colnames(cell_trans_data)
```

## Annotations in the data

* PTV: protein-truncating variants including nonsense, frameshift, and essential splice sites variants
* Missense 
* pLI: probability of loss of function intolerance 
* MPC: missense badness, PolyPhen-2, constraint score 

## Replicate Burden analysis for transmitted variants in Figure 1B 

```{r, echo=F}
trans_count=matrix(nrow=2, ncol=3)
rownames(trans_count)=c("transmitted", "untransmitted")
colnames(trans_count)=c("pLI>=0.995", "0.5<pLI<0.995", "MPC>=2")
###### pLI>=0.995
trans_count[1,1]=#cell_trans_data %>% filter(pLI>=0.995) %>% summarise(Transmitted_sibling) %>% sum(na.rm=T)+
cell_trans_data %>% filter(pLI>=0.995) %>% filter(PTV==T) %>% summarise(Transmitted_proband)%>%sum(na.rm=T) ### sum silbing and proband
trans_count[2,1]=#cell_trans_data %>% filter(pLI>=0.995) %>% summarise(Untransmitted_sibling) %>% sum(na.rm=T)+
cell_trans_data %>% filter(pLI>=0.995) %>% filter(PTV==T) %>% summarise(Untransmitted_proband)%>%sum(na.rm=T)
##### 0.5<pLI<0.95
trans_count[1,2]=#cell_trans_data %>% filter(pLI<0.995 & pLI>0.5) %>% summarise(Transmitted_sibling) %>% sum(na.rm=T)+
  cell_trans_data %>% filter(pLI<0.995 & pLI>0.5) %>% filter(PTV==T) %>% summarise(Transmitted_proband)%>%sum(na.rm=T)
trans_count[2,2]=#cell_trans_data %>% filter(pLI<0.995 & pLI>0.5) %>% summarise(Untransmitted_sibling) %>% sum(na.rm=T)+
  cell_trans_data %>% filter(pLI<0.995 & pLI>0.5) %>% filter(PTV==T) %>% summarise(Untransmitted_proband)%>%sum(na.rm=T)
###### MPC>=2
trans_count[1,3]=#cell_trans_data %>% filter(MPC>=2) %>% summarise(Transmitted_sibling) %>% sum(na.rm=T)+
  cell_trans_data %>% filter(MPC>=2) %>% summarise(Transmitted_proband)%>%sum(na.rm=T)
trans_count[2,3]=#cell_trans_data %>% filter(MPC>=2) %>% summarise(Untransmitted_sibling) %>% sum(na.rm=T)+
  cell_trans_data %>% filter(MPC>=2) %>% summarise(Untransmitted_proband)%>%sum(na.rm=T)
no.prob=5869; no.silb=1993

########## burden test
burden=matrix(nrow=3, ncol=3)
rownames(burden)=c("OR", "RR", "p.value")
colnames(burden)=c("pLI>=0.995", "0.5<pLI<0.995", "MPC>=2")
for (i in 1:ncol(trans_count))
  { 
  test=binom.test(trans_count[,i],alternative="greater")
  burden[1,i]=trans_count[1,i]/(no.prob-trans_count[1,i])/(trans_count[2,i]/(no.prob-trans_count[2,i]))
  burden[2,i]=trans_count[1,i]/trans_count[2,i]
  burden[3,i]=test$p.value
}
kable(trans_count, caption="variant count of transmitted PTV and MPC")
kable(burden, caption="Enrichment of transmited PTV and MPC in proband using binomial exact test")
```

In the paper, when $pLI>=0.995$ for PTV, there are 695 transmitted vs 557 untransmitted from 5869 parents, leading to RR 1.2 and p value 0.07?, and  these numbers are close to that in the above table. 

```{r, echo=F, eval=F}
# in Figure1 B. for PTV with pLI>=0.995, there are 695 transmitted vs 557 untransmitted in 5869 parents, thus 
RR=round(695/5869/(557/5869),1)
# how come the p value is 0.07? 
```

## Annotate the data by ANNOVAR

Since the data doesn't have allele frequency, use ANNOVAR to annotate it 

```{r, echo=T}
head(cell_trans_data)
```

### prepare the data in the appropriate format  for annovar 

To run annovar, the simplist format has 5 columns chr, star, end, ref, alt and each row is one variant. 

1.  use R to extract 5 required columns 

```{r, echo=T, eval=F}

cell_data=read.table("cell_trans_data.txt", header=T, fill=T) # note "fill" some lines with less than 38 elements
locus=matrix(cell_data$Variant)
split_locus=strsplit(locus, ":") # ":"  separates chr, start and end 
avinput=matrix(nrow=length(split_locus), ncol=5)  # format of avinput; chr, start, end, ref, alt
for (i in 1:length(split_locus))
  { 
   avinput[i,1:2]=split_locus[[i]][1:2]
   avinput[i,3]=split_locus[[i]][2]
   avinput[i,4:5]=split_locus[[i]][3:4]
}

#write.table(avinput, file="avinput.txt",row.names=F, col.names=F, quote=F)

```

2.  then transfer into tab-delimited columns

```{r, echo=T, eval=F}
head avinput.txt
1 861283 861283 G C
1 861287 861287 C T
1 861289 861289 G A
1 861341 861341 A G
1 861389 861389 C T
1 861398 861398 G A
1 865635 865635 G A
1 865645 865645 T G
1 865655 865655 T G
1 865686 865686 A G
```


```{r, echo=T, eval=F}
awk '{ for(i=1;i<=NF;i++){if(i==NF){printf("%s\n",$NF);}else {printf("%s\t",$i)}}}'
input.data > output.bed
```

3.  download relevant database 

```{r, echo=T, eval=F}
perl annotate_variation.pl -buildver hg19 -downdb -webfrom annovar exac03 humandb/
```


4.  run annovar function 

```{r, echo=T, eval=F}
perl annotate_variation.pl -filter -build hg19 -dbtype exac03 avinput humandb/
```


## Refine annotations 

Stratify six variant groups PTV (PLI>=0.995, 0.5-0.995, <0.5) and missense (MPC>=2, 1-2, 0-1) by AF (0.01-0.05, 0.001-0.01, <0.001), resulting in 18 categories. 


