---
title: "DS-ASD-constraint gene"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, echo=F}
# load mirage package
rm(list=ls()) # clean memory
set.seed(123)
library("devtools")
install_github('xinhe-lab/mirage')
library(mirage)
library(tidyverse)
```


```{r, echo=F}
loadRData <- function(fileName){
#loads an RData file, and returns it
    load(fileName)
    get(ls()[ls() != "fileName"])
}
```


```{r,echo=F, eval=F}
path1="C:\\Shengtong\\Research\\rare-var\\"
path2="C:\\Shengtong\\Research\\rare-var\\73416-DS-ASD\\"
All.Anno.Data=as_tibble(read.table(paste(path1, "AnnotatedTrans.txt", sep=""), header=T))
DS.ASD.sample=as_tibble(read.csv(paste(path2, "73416-DS-ASD.sample.csv", sep=""), header=T)[,-1])
gene.set=as.character(read.csv("C:/Shengtong/Research/rare-var/GeneSet/Samocha_2014NG/Samocha_2014NG_contraintgene.csv", header=T)$gene)
```

```{r, echo=F}
eight.partition=function(cand.data) # given gene data and annotations, do variant partitions
{
  par.evid=list()
  LoF.def=c("stopgain", "frameshift substitution", "splicing", "stoploss")
  par.evid[[1]]=which(cand.data$Annotation %in% LoF.def==T & cand.data$ExacAF<0.05 & cand.data$ExacAF>=0.01 )
  par.evid[[2]]=which(cand.data$Annotation %in% LoF.def==T &  cand.data$ExacAF<0.01)

  par.evid[[3]]=which(cand.data$Annotation %in% LoF.def==F & as.numeric(as.character(cand.data$Polyphen2.HDIV.score))>=0.957 & cand.data$ExacAF>=0.01 & cand.data$ExacAF<0.05)
  par.evid[[4]]=which(cand.data$Annotation %in% LoF.def==F & as.numeric(as.character(cand.data$Polyphen2.HDIV.score))>=0.957 & cand.data$ExacAF>=0.001 & cand.data$ExacAF<0.01)
  par.evid[[5]]=which(cand.data$Annotation %in% LoF.def==F & as.numeric(as.character(cand.data$Polyphen2.HDIV.score))>=0.957 &  cand.data$ExacAF<0.001)

  par.evid[[6]]=which(cand.data$Annotation %in% LoF.def==F & as.numeric(as.character(cand.data$Polyphen2.HDIV.score))<0.957 & cand.data$ExacAF>=0.01 & cand.data$ExacAF<0.05)
  par.evid[[7]]=which(cand.data$Annotation %in% LoF.def==F & as.numeric(as.character(cand.data$Polyphen2.HDIV.score))<0.957 & cand.data$ExacAF>=0.001 & cand.data$ExacAF<0.01)
  par.evid[[8]]=which(cand.data$Annotation %in% LoF.def==F & as.numeric(as.character(cand.data$Polyphen2.HDIV.score))<0.957 & cand.data$ExacAF<0.001)

  group.index=rep(NA, nrow(cand.data))
  for (i in 1:length(par.evid))
    group.index[par.evid[[i]]]=i
  gene.data=data.frame(ID=cand.data$ID, Gene=cand.data$Gene, No.case=cand.data$No.case, No.contr=cand.data$No.contr, group.index=group.index)
  gene.data=gene.data[complete.cases(gene.data),]
  return(gene.data)
}
```



```{r, echo=F, eval=F}
### this is for DS-ASD sample 
N1=2403; N0=2403
DS.ASD.sample[DS.ASD.sample =="."] <- NA
DS.ASD.sample$ExacAF[is.na(DS.ASD.sample$ExacAF)]=0 # set AF of NA to zero 
#Anno.Data=DS.ASD.sample[which(DS.ASD.sample$ExacAF<0.005 & DS.ASD.sample$Annotation!="synonymous SNV"),] # use AF cutoff and exclude synonumous SNV

Anno.Data=DS.ASD.sample[which(DS.ASD.sample$case.af<0.01 & DS.ASD.sample$control.af<0.01 & DS.ASD.sample$Annotation!="synonymous SNV"),] # use sample AF cutoff and exclude synonumous SNV

cand.data=as_tibble(Anno.Data[which(Anno.Data$Gene %in% gene.set),])
cand.data=cand.data %>% rename(Polyphen2.HDIV.score=Polyphen2.HDIV)
gene.data=eight.partition(cand.data) # partition variants into different groups 
ds.asd.result=mirage(gene.data, n1=2403, n2=2403, gamma=c(rep(6,5), rep(3,3)))

no.row=1+length(ds.asd.result$eta.est); no.col=2
parameter.estimate=matrix(nrow=no.row, ncol=no.col) # it may not have all 8 variant categories 
parameter.estimate[,1]=c(ds.asd.result$delta.est, ds.asd.result$eta.est)
parameter.estimate[,2]=c(ds.asd.result$delta.pvalue, ds.asd.result$eta.pvalue)
rownames(parameter.estimate)=c("delta", paste(names(ds.asd.result$eta.est), sep=""))
colnames(parameter.estimate)=c("estimate", "pval")
BF.PP.gene=ds.asd.result$BF.PP.gene
#save(parameter.estimate, BF.PP.gene, file="../output/para.est.constraint.gene.estimate.delta.DS.ASD.sampleAFcutoff1%.RData")
#############  run standalone mirage function 
#ds.asd.processed=pre.process(gene.data)
#actu.num.group=length(unique(ds.asd.processed$group.index))
#beta.init=runif(actu.num.group)
#ds.asd.result.processed=multi.group.func(ds.asd.processed, N1, N0, gamma.mean=3, #sigma=2, delta=0.2, beta.init, actu.num.group)
#############
```

```{r,echo=F, eval=F}
## check variant group index consistency before and after running mirage 
gene.id="SPEN"; var.id="chr1:16262483_16262484_G_C"
index.before=gene.data[gene.data$ID==var.id & gene.data$Gene==gene.id,]$group.index
gene.index=which(ds.asd.result$BF.PP.gene$Gene==gene.id)
gene.info=ds.asd.result$BF.all[[gene.index]]
index.after=gene.info[gene.info$ID==var.id,]$original.index
index.before==index.after
```




```{r, echo=F, eval=F}
# this is for old sample 
N1=4315; N0=4315
All.Anno.Data[All.Anno.Data =="."] <- NA
All.Anno.Data$ExacAF[is.na(All.Anno.Data$ExacAF)]=0 # set AF of NA to zero
old.data=All.Anno.Data[which(All.Anno.Data$ExacAF<0.05 & All.Anno.Data$Annotation!="synonymous SNV"),] # use AF cutoff and exclude synonumous SNV
Anno.vart.set=as.character(old.data$ID[which(old.data$Gene %in% gene.set)])
Anno.cand.data=old.data[which(old.data$ID %in% Anno.vart.set),]
Anno.gene.data=eight.partition(Anno.cand.data)
old.result=mirage(Anno.gene.data, n1=4315, n2=4315, gamma=c(rep(6,5), rep(3,3)))

################# use mirage standalone function not package to see if we are able to have identical result 
pre.process=function(Anno.gene.data) # pre-process the data to run function-multi.group.func
{
    overlap.id=Anno.gene.data$ID
    overlap.data=Anno.gene.data[Anno.gene.data$ID %in% overlap.id,]
    order.overlap.data=overlap.data[order(overlap.data$group.index, decreasing=F),]
    psbl.index=unique(order.overlap.data$group.index); actu.num.group=length(psbl.index)
    delta.init=runif(1); #beta.init=runif(actu.num.group)
    beta.init=rep(0.1, actu.num.group)
    for (j in 1:actu.num.group)
      order.overlap.data$group.index[order.overlap.data$group.index==psbl.index[j]]=j # re-index the group labels
  #  delta=runif(1)
    return(order.overlap.data)
}

para.est=multi.group.func(order.overlap.data, N1, N0, gamma.mean=3, sigma=2, delta=0.1, beta.init, actu.num.group)
```



## Parameter estimate 

### when AF < 5\%

$\widehat{\delta}_{old}=0.18$ and $\widehat{\delta}_{DS-ASD}=0.27$


```{r, echo=F}

parameter.estimate=tibble(cate=rep(c("LoF; 1%<AF<5%", "LoF; AF<1%", "Damaging; 1%<AF<5%", "Damaging; 0.1%<AF<1%", "Damaging; AF<0.1%", "Non-damaging; 1%<AF<5%", "Non-damaging; 0.1%<AF<1%", "Non-damaging; AF<0.1%"),2), group=rep(c("old", "DS-ASD"),each=8), estimate=c(0, 0.41, 0, 0, 0.16, 0, 0, 0.17, 1, 0.07, 0.21, 0.02, 0.03, 0.31, 0.16, 0.11) )

ggplot(data=parameter.estimate, aes(x=cate, y=estimate, fill=group)) +
geom_bar(stat="identity", position=position_dodge())+
   xlab("")+
   ylab(expression(paste(hat(eta))))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=15))
```


### when AF < 1\%


$\widehat{\delta}_{old}=0.18$ and $\widehat{\delta}_{DS-ASD}=0.22$


```{r, echo=F}
parameter.estimate=tibble(cate=rep(c("LoF; 1%<AF<5%", "LoF; AF<1%", "Damaging; 1%<AF<5%", "Damaging; 0.1%<AF<1%", "Damaging; AF<0.1%", "Non-damaging; 1%<AF<5%", "Non-damaging; 0.1%<AF<1%", "Non-damaging; AF<0.1%"),2), group=rep(c("old", "DS-ASD"),each=8), estimate=c(0, 0.41, 0, 0, 0.16, 0, 0, 0.17, 0, 0.08, 0, 0.03, 0.04, 0, 0.19, 0.12) )

ggplot(data=parameter.estimate, aes(x=cate, y=estimate, fill=group)) +
geom_bar(stat="identity", position=position_dodge())+
   xlab("")+
   ylab(expression(paste(hat(eta))))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=15))
```


### when AF < 0.5\%


$\widehat{\delta}_{old}=0.18$ and $\widehat{\delta}_{DS-ASD}=0.17$


```{r, echo=F}
parameter.estimate=tibble(cate=rep(c("LoF; 1%<AF<5%", "LoF; AF<1%", "Damaging; 1%<AF<5%", "Damaging; 0.1%<AF<1%", "Damaging; AF<0.1%", "Non-damaging; 1%<AF<5%", "Non-damaging; 0.1%<AF<1%", "Non-damaging; AF<0.1%"),2), group=rep(c("old", "DS-ASD"),each=8), estimate=c(0, 0.41, 0, 0, 0.16, 0, 0, 0.17, 0, 0.06, 0, 0.02, 0.05, 0, 0.21, 0.15) )

ggplot(data=parameter.estimate, aes(x=cate, y=estimate, fill=group)) +
geom_bar(stat="identity", position=position_dodge())+
   xlab("")+
   ylab(expression(paste(hat(eta))))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=15))
```


### when sample (case or control) AF < 5\%

$\widehat{\delta}_{old}=0.18$ and $\widehat{\delta}_{DS-ASD}=0.29$


```{r, echo=F}

parameter.estimate=tibble(cate=rep(c("LoF; 1%<AF<5%", "LoF; AF<1%", "Damaging; 1%<AF<5%", "Damaging; 0.1%<AF<1%", "Damaging; AF<0.1%", "Non-damaging; 1%<AF<5%", "Non-damaging; 0.1%<AF<1%", "Non-damaging; AF<0.1%"),2), group=rep(c("old", "DS-ASD"),each=8), estimate=c(0, 0.41, 0, 0, 0.16, 0, 0, 0.17, 1, 0.06, 0.19, 0.02, 0.01, 0.21, 0.14, 0.09) )

ggplot(data=parameter.estimate, aes(x=cate, y=estimate, fill=group)) +
geom_bar(stat="identity", position=position_dodge())+
   xlab("")+
   ylab(expression(paste(hat(eta))))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=15))
```

### when sample (case or control) AF < 1\%

$\widehat{\delta}_{old}=0.18$ and $\widehat{\delta}_{DS-ASD}=0.14$


```{r, echo=F}

parameter.estimate=tibble(cate=rep(c("LoF; 1%<AF<5%", "LoF; AF<1%", "Damaging; 1%<AF<5%", "Damaging; 0.1%<AF<1%", "Damaging; AF<0.1%", "Non-damaging; 1%<AF<5%", "Non-damaging; 0.1%<AF<1%", "Non-damaging; AF<0.1%"),2), group=rep(c("old", "DS-ASD"),each=8), estimate=c(0, 0.41, 0, 0, 0.16, 0, 0, 0.17, 0, 0, 0, 0, 0, 0.32, 0.15, 0.04) )

ggplot(data=parameter.estimate, aes(x=cate, y=estimate, fill=group)) +
geom_bar(stat="identity", position=position_dodge())+
   xlab("")+
   ylab(expression(paste(hat(eta))))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=10))+
theme(text = element_text(size=15))
```



## Top genes 

### when AF<5\%

```{r, echo=F}
rm("parameter.estimate", "BF.PP.gene")
#const.gene.BF.DS.ASD=loadRData("../output/para.est.constraint.gene.estimate.delta.DS.ASD.RData") loadRData only load one item in RData
load("../output/para.est.constraint.gene.estimate.delta.DS.ASD.RData") # load all items stored in RData
BF.PP.gene=as_tibble(BF.PP.gene)
BF.PP.gene.sort=arrange(BF.PP.gene, desc(post.prob))
BF.PP.gene.sort


```

### when AF < 1\%

There are 16 genes with posterior probability of 1.

```{r, echo=F}
rm("parameter.estimate", "BF.PP.gene")
#const.gene.BF.DS.ASD=loadRData("../output/para.est.constraint.gene.estimate.delta.DS.ASD.RData") loadRData only load one item in RData
load("../output/para.est.constraint.gene.estimate.delta.DS.ASD.AFcutoff1%.RData") # load all items stored in RData
BF.PP.gene=as_tibble(BF.PP.gene)
BF.PP.gene.sort=arrange(BF.PP.gene, desc(post.prob))
BF.PP.gene.sort

```


### when AF < 0.5\%


There are 16 genes with posterior probability of 1.

```{r, echo=F}
rm("parameter.estimate", "BF.PP.gene")
#const.gene.BF.DS.ASD=loadRData("../output/para.est.constraint.gene.estimate.delta.DS.ASD.RData") loadRData only load one item in RData
load("../output/para.est.constraint.gene.estimate.delta.DS.ASD.AFcutoff0.5%.RData") # load all items stored in RData
BF.PP.gene=as_tibble(BF.PP.gene)
BF.PP.gene.sort=arrange(BF.PP.gene, desc(post.prob))
BF.PP.gene.sort

```


### when sample AF<5%

```{r, echo=F}
rm("parameter.estimate", "BF.PP.gene")
#const.gene.BF.DS.ASD=loadRData("../output/para.est.constraint.gene.estimate.delta.DS.ASD.RData") loadRData only load one item in RData
load("../output/para.est.constraint.gene.estimate.delta.DS.ASD.sampleAFcutoff5%.RData") # load all items stored in RData
BF.PP.gene=as_tibble(BF.PP.gene)
BF.PP.gene.sort=arrange(BF.PP.gene, desc(post.prob))
BF.PP.gene.sort
```


### when sample AF<1%

```{r, echo=F}
rm("parameter.estimate", "BF.PP.gene")
#const.gene.BF.DS.ASD=loadRData("../output/para.est.constraint.gene.estimate.delta.DS.ASD.RData") loadRData only load one item in RData
load("../output/para.est.constraint.gene.estimate.delta.DS.ASD.sampleAFcutoff1%.RData") # load all items stored in RData
old.top.genes=c("CYFIP1", "EP400", "FBN1", "SRCAP", "DYNC1H1", "ABCA2", "DOCK4", "CACNA1D", "CHD8", "TRIP12")
BF.PP.gene=as_tibble(BF.PP.gene)
BF.PP.gene.sort=arrange(BF.PP.gene, desc(post.prob))
BF.PP.gene.sort
BF.PP.gene.sort %>% filter(Gene%in%old.top.genes)
```


## Gene SPEN when AF<1%

there are two variants driving the signal 
```{r, echo=T, eval=F}
                             ID Gene No.case No.contr category       var.BF original.index
1321 chr1:16262483_16262484_G_C SPEN      30      204        4 2.065188e+25              7
1322 chr1:16262484_16262485_T_C SPEN      28      191        4 2.509197e+23              7
```


## Gene CDC27 when AF<5%

This gene has a very high BF driving by many variants with big BF, such as 
`stopgain` variant at `chr17:45234411_45234412_G_T` has 28/217 variants in cases and control, leading to BF of  3.638739e+20. 




```{r, echo=F, eval=F}

gene.index=which(BF.PP.gene$Gene=="CDC27")
gene.cdc27=ds.asd.result$BF.all[[gene.index]]
head(gene.cdc27)
```










