---
title: "simulation logistic prior"
output: html_document
---

```{r, echo=F, message=F, warning=F}
library(tidyverse)
```

## data generation process 




## one Annotation group 

Suppose there is one annotation group affecting the risk of variants via logistic function 

```{r, echo=F, message=F, warning=F}
beta_estimate=c(2.016002, 1.998177, 1.994396, 2.024530, 1.988762, 1.989376, 1.995133, 2.011443, 1.977110, 1.969553) # this estimate is obtained by running code: LogisticPriorOneAnnoFeature.R 


beta_estimate_summary=data.frame(estimate=mean(beta_estimate), lower=mean(beta_estimate)-sd(beta_estimate), upper=mean(beta_estimate)+sd(beta_estimate), var="effect size: beta")
ggplot(beta_estimate_summary, aes(x=var, y=estimate, fill=var)) + # draw the OR plot 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_hline(yintercept = 2,linetype="dashed", 
                color = "red", size=1)+
  xlab("")+
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.2,
                 position=position_dodge(.9)) 

```

* run 10 replicates 
* one annotation group with true $\beta=2$



## multiple annotation groups 

```{r, echo=F, message=F, warning=F}
beta_estimate=c(0.4822893, 2.000852, 0.4909131, 1.976159, 0.5044987, 2.000259, 0.4988077, 1.980916, 0.4992905, 2.009325, 0.4908483,  2.003779, 0.5031419, 2.015141,0.5085360, 2.005088,0.5025723, 1.995207, 0.4998137, 2.019582) 
# this estimate is obtained by running code:C:\Shengtong\Research\rare-var\RareVariant\rare-var-project\code\LogisticPriorManyAnnoFeature.R 

beta1_estimate_index=c(1,3,5,7,9,11,13,15,17,19)
beta2_estimate_index=beta1_estimate_index+1

beta_estimate_summary=data.frame(
  estimate=c(mean(beta_estimate[beta1_estimate_index]), mean(beta_estimate[beta2_estimate_index])),  lower=c(mean(beta_estimate[beta1_estimate_index])-sd(beta_estimate[beta1_estimate_index]), mean(beta_estimate[beta2_estimate_index])-sd(beta_estimate[beta2_estimate_index])), upper=c(mean(beta_estimate[beta1_estimate_index])+sd(beta_estimate[beta1_estimate_index]), mean(beta_estimate[beta2_estimate_index])+sd(beta_estimate[beta2_estimate_index])), var=c("beta1", "beta2"))

ggplot(beta_estimate_summary, aes(x=var, y=estimate, fill=var)) + # draw the OR plot 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
 # geom_hline(yintercept = 2,linetype="dashed", 
  #              color = "red", size=1)+
  xlab("")+
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.2,
                 position=position_dodge(.9)) 

```

* covariates $A_1, A_2$ are binary, and they need to be non-correlated as much as possible, otherwise $\widehat{\beta}_1, \widehat{\beta}_2$ will be messed up.  

* 10 repetitions 

* true $\beta_1=0.5, \beta_2=2$


## more power of logistic prior over disjoint group prior 



